QPLIB: A Library of Quadratic Programming Instances

### 2018-02-05 QPLIB-3.tex
Update QPLIB-3.tex

The number of positive (i.e., larger than $10^{-12}$) and negative (i.e., smaller than $-10^{-12}$) eigenvalues of $Q^0$ allowed us to identify the
objective function type, as in presence of at least one negative (positive) eigenvalue
the objective function is nonconvex (nonconcave). Finally, as far as the constraint types
are concerned, we collected the following information:
%
\begin{itemize}
 \item the number of linear constraints, % (\emph{\# lin}),
 %
 \item the number of quadratic constraints, % (\emph{\# quad}),
 %
 \item the number of convex constraints, and % (\emph{\# conv}),
 %
 \item the number of variable bounds (for non-binary variables). %(\emph{\# box}).
\end{itemize}
%
A constraint is considered quadratic if it contains at least one nonzero in
a quadratic term (if present). Among the quadratic constraints, the ones whose
Hessians have only non-negative eigenvalues (when $c_u^i < \infty$) and
and non-positive eigenvalues (when $c_l^i > - \infty$)
are classified as convex constraints; thus, a quadratic constraint with
two sided, finite bounds is non-convex.
Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
 All this information allowed us to analyse the gathered instances and to perform the
filters described in the the next paragraph.

@@

During the development of the library, a discussion ensued about
the expected goals that we wished to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.

>>>>

The number of positive (i.e., larger than $10^{-12}$) and negative (i.e., smaller than $-10^{-12}$) eigenvalues of $Q^0$ allowed us to identify the
objective function type, as in presence of at least one negative (positive) eigenvalue
the objective function is nonconvex (nonconcave). Finally, as far as the constraint types
are concerned, we collected the following information:
%
\begin{itemize}
 \item the number of linear constraints, % (\emph{\# lin}),
 %
 \item the number of quadratic constraints, % (\emph{\# quad}),
 %
 \item the number of convex constraints, and % (\emph{\# conv}),
 %
 \item the number of variable bounds (for non-binary variables). %(\emph{\# box}).
\end{itemize}
%
A constraint is considered quadratic if it contains at least one nonzero in
a quadratic term (if present). Among the quadratic constraints, the ones whose
Hessians have only non-negative eigenvalues (when $c_u^i < \infty$) and
and non-positive eigenvalues (when $c_l^i > - \infty$)
are classified as convex constraints; thus, a quadratic constraint with
two sided, finite bounds is non-convex.
Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex; however, only some solvers are capable of properly exploiting this property.
 All this information allowed us to analyse the gathered instances and to perform the
filters described in the the next paragraph.

@@

During the development of the library, a discussion ensued about
the expected goals that we wished to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.
